An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-22 VLDB2026/main.tex
Update main.tex

	\textit{Single-Thread Search.}  
	\textcolor{violet}{As shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}, under near-perfect recall (recall approaching 1) conditions,} search often requires traversing larger candidate sets or deeper graph paths. This leads to increased computational cost and a sharp QPS \sout{(queries per second)} drop for most methods. Despite this, UNG maintains high QPS due to its LNG structure, which eliminates unnecessary online filtering of irrelevant attributes, thereby improving query efficiency. \textcolor{violet}{Puck and CAPS perform relatively well on some datasets. Partition-based methods have limitations regarding different vector data distributions, and how to configure different partitions for different datasets remains a challenge. However, this also demonstrates that IVF-based methods are still very promising for hybrid queries.}

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation.  CAPS shows a slight performance decline \underline{due to the increased query overhead caused by excessive} attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases \underline{because data points may duplicate across} \underline{multiple subgraphs}. 3) \underline{The fusion distance of NHQ exhibits sensitivity} \underline{to the number of attributes}, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, degenerating into a situation closer to inverted retrieval, and the effect may be greatly impaired.

@@

	As illustrated in Figure~\ref{fig:exp_3_1}, the evaluated algorithms exhibit varying degrees of sensitivity to attribute distribution shifts. \textcolor{cyan}{The proximity of the four curves (representing different attribute distributions) in the figure reflects the algorithm's sensitivity to distributional variations. When the curves converge closely, the algorithm demonstrates strong resilience against attribute distribution shifts. Conversely, greater divergence among the curves indicates higher distributional dependence. Accordingly, the "Strong" label signifies robust performance consistency across varying distributions, while "Weak" denotes pronounced performance fluctuations under distributional changes.} Specifically, ACORN, Milvus, Puck, CAPS, StitchedVamana, and UNG demonstrate strong robustness to such shifts. In contrast, Faiss, FilteredVamana, NHQ, VBASE, and PASE show notable sensitivity.

@@

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets.}}}

@@

>>>>

	\textit{Single-Thread Search.}  
	\textcolor{violet}{As shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}, under near-perfect recall (recall approaching 1) conditions,} search often requires traversing larger candidate sets or deeper graph paths. This leads to increased computational cost and a sharp QPS (queries per second) drop for most methods. Despite this, UNG maintains high QPS due to its LNG structure, which eliminates unnecessary online filtering of irrelevant attributes, thereby improving query efficiency. \textcolor{violet}{Puck and CAPS perform relatively well on some datasets. Partition-based methods have limitations regarding different vector data distributions, and how to configure different partitions for different datasets remains a challenge. However, this also demonstrates that IVF-based methods are still very promising for hybrid queries.}

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation.  CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, degenerating into a situation closer to inverted retrieval, and the effect may be greatly impaired.

@@

	As illustrated in Figure~\ref{fig:exp_3_1}, the evaluated algorithms exhibit varying degrees of sensitivity to attribute distribution shifts. \textcolor{cyan}{The closer the four curves (corresponding to the four distributions) in the figure approach each other, the less the algorithm is affected by attribute distributions. Conversely, the farther they diverge, the more significant the impact of attribute distributions on the algorithm becomes.} Specifically, ACORN, Milvus, Puck, CAPS, StitchedVamana, and UNG demonstrate strong robustness to such shifts. In contrast, Faiss, FilteredVamana, NHQ, VBASE, and PASE show notable sensitivity.

@@

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets.}}} 
	\textcolor{violet}{On the 100M dataset, UNG and NHQ still exhibit excellent performance, demonstrating their good adaptability to large-scale datasets; however, this also comes with significant memory overhead. The clustering-based algorithm \texttt{Faiss+HQI\_Batch} shows moderate performance but has very low memory overhead, which makes it highly attractive for practical applications. We did not present the experimental results for StitchedVamana because we found that its recall consistently remained at a very low level. Its search starting point is set to 0 by default, and when the entry point is far from the target area, it easily gets stuck in a local optimum, making it difficult to navigate to the vicinity of the correct point.}

@@

	\begin{table}[t]
		\centering
		\caption{Algorithm Recommendation per Scenario}
		% 移除 \resizebox 和它对应的外层大括号
		% 现在字体大小命令直接作用于 tabular 环境
		\fontsize{5pt}{6pt}\selectfont % 字体大小设置为 5pt
		\begin{tabular}{|c|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{CT} & \textbf{IS} & \textbf{BPM} & \textbf{SPM}\\

			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
			ACORN-1 & Puck, UNG & iRange & iRange & iRange\\[1pt]
			\hline
		\end{tabular}
		\label{tab:algo_scenarios}
	\end{table}
